---
title: "CE Data and Passenger Data"
format: html
---

# Load Required Libraries

```{r}
library(readr)
library(readxl)
library(dplyr)
library(lubridate)
library(ggplot2)
library(scales)
library(tidyr)
```

# Import Passenger and Event Data

```{r}
passengers <- read_csv("pnfs_passengers_long.csv", show_col_types = FALSE)
events <- read_excel("Events Tracker all - Copy - Copy.xlsx")
```

# Inspect Event Data Structure

```{r}
colnames(events)
head(events)
```

# Clean Event Dates and Create Year/Month

```{r}
events_clean <- events %>%
  mutate(
    date = as.Date(`Event Date`),
    year = year(date),
    month = month(date)
  )
```

# Check Event Date Quality

```{r}
summary(events_clean$date)
```

# Aggregate Events by Month

```{r}
events_monthly <- events_clean %>%
  group_by(year, month) %>%
  summarise(
    total_events = n(),
    .groups = "drop"
  )

print(events_monthly)

```

Counts how many community engagement events happened in each month.

This creates a monthly event activity table.

# Aggregate Passenger Data by Airline and Month

```{r}
passengers_airline <- passengers %>%
  mutate(
    date = as.Date(date),
    year = year(date),
    month = month(date)
  ) %>%
  group_by(year, month, airline) %>%
  summarise(
    passengers = sum(passengers, na.rm = TRUE),
    .groups = "drop"
  )

print(passengers_airline)
```

Converts passenger dates and summarizes passengers by:

year
month
airline

This creates the airline-level traffic dataset.

# Merge Event Data with Passenger Data

```{r}
data_airline <- passengers_airline %>%
  left_join(events_monthly, by = c("year", "month")) %>%
  mutate(
    total_events = ifelse(is.na(total_events), 0, total_events),
    season = case_when(
      month %in% c(12, 1, 2) ~ "Winter",
      month %in% c(3, 4, 5) ~ "Spring",
      month %in% c(6, 7, 8) ~ "Summer",
      month %in% c(9, 10, 11) ~ "Fall"
    ),
    season = factor(season, levels = c("Winter", "Spring", "Summer", "Fall"))
  ) %>%
  arrange(year, month, airline)

print(data_airline)
```

Merges monthly event totals into the airline passenger dataset.
It also:

fills missing event months with 0
creates a season variable

This is your main analysis dataset.

# Summarize by Season and Airline

```{r}
season_summary <- data_airline %>%
  group_by(season) %>%
  summarise(
    avg_passengers = mean(passengers, na.rm = TRUE),
    avg_events = mean(total_events, na.rm = TRUE),
    .groups = "drop"
  )

print(season_summary)

airline_summary <- data_airline %>%
  group_by(airline) %>%
  summarise(
    total_passengers = sum(passengers, na.rm = TRUE),
    avg_passengers = mean(passengers, na.rm = TRUE),
    avg_events = mean(total_events, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(total_passengers))

print(airline_summary)
```

Creates two summaries:

season_summary = average passengers and events by season
airline_summary = total and average passengers by airline

This gives context before modeling.

# Measure Airline-Level Correlation


```{r}
airline_effect <- data_airline %>%
  group_by(airline) %>%
  summarise(
    correlation = cor(passengers, total_events, use = "complete.obs"),
    .groups = "drop"
  ) %>%
  arrange(desc(correlation))

print(airline_effect)
```

Calculates the correlation between event activity and passengers for each airline.

This helps identify which airlines appear more responsive to community engagement.

# Build Base Regression Model


```{r}
model_base <- lm(passengers ~ total_events + airline + season, data = data_airline)
summary(model_base)
```

Builds a linear regression model to estimate passenger traffic using:

number of events
airline
season

This is your baseline relationship model.

# Build Interaction Model

```{r}
model_interaction <- lm(passengers ~ total_events * airline + season, data = data_airline)
summary(model_interaction)
```

Builds a stronger model by allowing the effect of total_events to vary by airline.

This helps answer:

Do some airlines benefit more from events than others?

# Choose Final Simulation Model


```{r}
model_sim <- model_interaction
```

Sets the interaction model as the main model used for scenario prediction.

# Predict Baseline Passenger Levels

```{r}
data_airline <- data_airline %>%
  mutate(
    baseline_pred = predict(model_sim, newdata = data_airline)
  )
```

Uses the model to estimate baseline predicted passengers for each airline-month record.

This becomes the comparison point for all scenarios.

# Simulate Summer +20% Event Scenario

```{r}
scenario_summer <- data_airline %>%
  mutate(
    simulated_events = ifelse(season == "Summer", total_events * 1.20, total_events)
  )

scenario_summer$scenario_pred <- predict(
  model_sim,
  newdata = scenario_summer %>% mutate(total_events = simulated_events)
)

scenario_summer <- scenario_summer %>%
  mutate(
    incremental_passengers = scenario_pred - baseline_pred,
    scenario = "Summer +20%"
  )

```

Simulates a 20% increase in events during summer only, then predicts the passenger impact.

# Simulate Winter +20% Event Scenario

```{r}
scenario_winter <- data_airline %>%
  mutate(
    simulated_events = ifelse(season == "Winter", total_events * 1.20, total_events)
  )

scenario_winter$scenario_pred <- predict(
  model_sim,
  newdata = scenario_winter %>% mutate(total_events = simulated_events)
)

scenario_winter <- scenario_winter %>%
  mutate(
    incremental_passengers = scenario_pred - baseline_pred,
    scenario = "Winter +20%"
  )
```

Same process as above, but increases winter events by 20%.

# Simulate Fall +20% Event Scenario


```{r}
scenario_fall <- data_airline %>%
  mutate(
    simulated_events = ifelse(season == "Fall", total_events * 1.20, total_events)
  )

scenario_fall$scenario_pred <- predict(
  model_sim,
  newdata = scenario_fall %>% mutate(total_events = simulated_events)
)

scenario_fall <- scenario_fall %>%
  mutate(
    incremental_passengers = scenario_pred - baseline_pred,
    scenario = "Fall +20%"
  )
```

Tests the effect of increasing fall events by 20%.

# Simulate Spring +20% Event Scenario

```{r}
scenario_spring <- data_airline %>%
  mutate(
    simulated_events = ifelse(season == "Spring", total_events * 1.20, total_events)
  )

scenario_spring$scenario_pred <- predict(
  model_sim,
  newdata = scenario_spring %>% mutate(total_events = simulated_events)
)

scenario_spring <- scenario_spring %>%
  mutate(
    incremental_passengers = scenario_pred - baseline_pred,
    scenario = "Spring +20%"
  )
```

Tests the effect of increasing spring events by 20%.

# Combine Seasonal Scenario Results

```{r}
scenario_results <- bind_rows(
  scenario_summer,
  scenario_winter,
  scenario_fall,
  scenario_spring
)
```

Combines all seasonal scenarios into one table for comparison.

# Summarize Seasonal Scenario Impact

```{r}
scenario_summary <- scenario_results %>%
  group_by(scenario) %>%
  summarise(
    total_incremental_passengers = sum(incremental_passengers, na.rm = TRUE),
    avg_incremental_passengers = mean(incremental_passengers, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(total_incremental_passengers))

print(scenario_summary)
```

Shows how each airline performs under each scenario.

# Identify Best Airlines by Scenario

```{r}
scenario_airline_summary <- scenario_results %>%
  group_by(scenario, airline) %>%
  summarise(
    total_incremental_passengers = sum(incremental_passengers, na.rm = TRUE),
    avg_incremental_passengers = mean(incremental_passengers, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(scenario, desc(total_incremental_passengers))

print(scenario_airline_summary)
```

Finds the top 5 airlines with the highest projected passenger gains under each scenario.

# Plot Events vs Passengers by Airline

```{r}
best_airlines_by_scenario <- scenario_airline_summary %>%
  group_by(scenario) %>%
  slice_max(order_by = total_incremental_passengers, n = 5) %>%
  ungroup()

print(best_airlines_by_scenario)
```



```{r}
ggplot(data_airline, aes(x = total_events, y = passengers, color = airline)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Community Engagement Events vs Passengers by Airline",
    x = "Total Events",
    y = "Passengers"
  ) +
  theme_minimal()
```


```{r}
ggplot(scenario_summary, aes(x = reorder(scenario, total_incremental_passengers), y = total_incremental_passengers)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Incremental Passengers by Seasonal Event Scenario",
    x = "Scenario",
    y = "Incremental Passengers"
  ) +
  theme_minimal()

```


```{r}
ggplot(best_airlines_by_scenario, aes(x = reorder(airline, total_incremental_passengers), y = total_incremental_passengers)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~scenario, scales = "free_y") +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Top Airlines by Seasonal Event Scenario",
    x = "Airline",
    y = "Incremental Passengers"
  ) +
  theme_minimal()
```



```{r}
value_per_passenger <- 20

scenario_summary <- scenario_summary %>%
  mutate(
    incremental_value = total_incremental_passengers * value_per_passenger
  )

print(scenario_summary)
```


```{r}
scenarios <- data.frame(
  scenario = c("Baseline", "+10%", "+20%", "+30%"),
  multiplier = c(1.00, 1.10, 1.20, 1.30)
)
```


```{r}
scenario_results <- lapply(1:nrow(scenarios), function(i) {

  mult <- scenarios$multiplier[i]
  scen_name <- scenarios$scenario[i]

  temp <- data_airline %>%
    mutate(
      simulated_events = total_events * mult
    )

  temp$predicted <- predict(
    model_sim,
    newdata = temp %>% mutate(total_events = simulated_events)
  )

  temp <- temp %>%
    mutate(
      baseline = predict(model_sim, newdata = data_airline),
      incremental_passengers = predicted - baseline,
      scenario = scen_name
    )

  return(temp)
})

scenario_results <- bind_rows(scenario_results)
```


```{r}
scenario_summary <- scenario_results %>%
  group_by(scenario) %>%
  summarise(
    total_incremental_passengers = sum(incremental_passengers, na.rm = TRUE),
    avg_incremental_passengers = mean(incremental_passengers, na.rm = TRUE),
    .groups = "drop"
  )

print(scenario_summary)
```

```{r}
event_cost <- 114958
value_per_passenger <- 20

scenario_summary <- scenario_summary %>%
  mutate(
    incremental_value = total_incremental_passengers * value_per_passenger,
    roi = (incremental_value - event_cost) / event_cost
  )

print(scenario_summary)
```


```{r}
best_scenario <- scenario_summary %>%
  arrange(desc(roi))

print(best_scenario)
```


```{r}
ggplot(scenario_summary, aes(x = scenario, y = roi)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "ROI by Engagement Scenario",
    x = "Scenario",
    y = "ROI"
  ) +
  theme_minimal()
```

```{r}
print(scenario_summary$roi)
```


```{r}
ggplot(scenario_summary, aes(x = scenario, y = total_incremental_passengers)) +
  geom_col() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Incremental Passengers by Scenario",
    x = "Scenario",
    y = "Passengers"
  ) +
  theme_minimal()
```


```{r}
event_cost <- 114958
value_per_passenger <- 20

scenario_summary <- scenario_summary %>%
  mutate(
    incremental_value = total_incremental_passengers * value_per_passenger,
    roi_proxy = (incremental_value - event_cost) / event_cost,
    return_multiple = incremental_value / event_cost
  )

print(scenario_summary)
```

```{r}
ggplot(scenario_summary, aes(x = scenario, y = roi_proxy)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "ROI Proxy by Engagement Scenario",
    x = "Scenario",
    y = "ROI Proxy"
  ) +
  theme_minimal()
```

```{r}
ggplot(scenario_summary, aes(x = scenario, y = return_multiple)) +
  geom_col() +
  labs(
    title = "Economic Return Multiple by Engagement Scenario",
    x = "Scenario",
    y = "Return Multiple (x)"
  ) +
  theme_minimal()
```

```{r}
scenario_summary <- scenario_summary %>%
  mutate(
    roi_display = roi_proxy / 100
  )
```



```{r}
ggplot(scenario_summary, aes(x = scenario, y = roi_display)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "ROI by Engagement Scenario",
    x = "Scenario",
    y = "ROI"
  ) +
  theme_minimal()
```


```{r}
scenario_summary <- scenario_summary %>%
  mutate(
    incremental_value = total_incremental_passengers * value_per_passenger,
    roi_proxy = (incremental_value - event_cost) / event_cost,
    return_multiple = incremental_value / event_cost
  ) %>%
  select(scenario, total_incremental_passengers, incremental_value, roi_proxy, return_multiple)

print(scenario_summary)
```

