---
title: "Ontario Airport A04"
author: "Saurabh Parate"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
    code-fold: show
    code-summary: "Show code"
    df-print: paged
execute:
  echo: true
  warning: false
  message: false
  error: false
---

#testing
# Community Engagement Tracker Data Analysis

## Install / load packages

```{r}
library(tidyverse)
library(lubridate)
library(scales)
library(readr)
library(dplyr)
```

## Read the cleaned data

```{r}
passengers <- read_csv("pnfs_passengers_long.csv", show_col_types = FALSE)
metrics <- read_csv("pnfs_key_metrics_long.csv", show_col_types = FALSE)
```

## Inspect the data

```{r}
glimpse(passengers)
glimpse(metrics)

summary(passengers$passengers)
summary(metrics$value)
```

## Basic cleaning

```{r}
passengers <- passengers %>%
  mutate(
    date = as.Date(date),
    year = as.integer(year),
    month_num = as.integer(month_num),
    passengers = as.numeric(passengers)
  ) %>%
  filter(!is.na(passengers))

metrics <- metrics %>%
  mutate(
    date = as.Date(date),
    year = as.integer(year),
    month_num = as.integer(month_num),
    value = as.numeric(value)
  ) %>%
  filter(!is.na(value))
```

## Total passengers by month

```{r}
monthly_total <- passengers %>%
  group_by(date, year, month_num) %>%
  summarise(total_passengers = sum(passengers, na.rm = TRUE), .groups = "drop") %>%
  arrange(date)

print(monthly_total)
```

## Total passengers by airline

```{r}
airline_totals <- passengers %>%
  group_by(airline) %>%
  summarise(
    total_passengers = sum(passengers, na.rm = TRUE),
    avg_monthly_passengers = mean(passengers, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(total_passengers))

print(airline_totals)
```

## Top 10 airlines

```{r}
top_10_airlines <- airline_totals %>%
  slice_max(order_by = total_passengers, n = 10)

print(top_10_airlines)
```

## Yearly passenger trend

```{r}
yearly_total <- passengers %>%
  group_by(year) %>%
  summarise(total_passengers = sum(passengers, na.rm = TRUE), .groups = "drop") %>%
  arrange(year)

print(yearly_total)
```

## Year-over-year monthly change

```{r}
monthly_yoy <- monthly_total %>%
  arrange(month_num, year) %>%
  group_by(month_num) %>%
  mutate(
    yoy_change = total_passengers - lag(total_passengers),
    yoy_pct = (total_passengers - lag(total_passengers)) / lag(total_passengers) * 100
  ) %>%
  ungroup()

print(monthly_yoy)
```

## Visualizations

### Total passengers over time

```{r}
ggplot(monthly_total, aes(x = date, y = total_passengers)) +
  geom_line(linewidth = 1) +
  geom_point() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Ontario Airport Monthly Passenger Trend",
    x = "Date",
    y = "Total Passengers"
  ) +
  theme_minimal()
```

### Top airlines

```{r}
top_airlines_data <- passengers %>%
  filter(airline %in% top_10_airlines$airline) %>%
  group_by(airline) %>%
  summarise(total_passengers = sum(passengers, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(total_passengers))

ggplot(top_airlines_data, aes(x = reorder(airline, total_passengers), y = total_passengers)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Top 10 Airlines by Total Passengers",
    x = "Airline",
    y = "Passengers"
  ) +
  theme_minimal()

```

### Airline trend over time for top 5 airlines

```{r}
top_5 <- top_10_airlines %>% slice_head(n = 5) %>% pull(airline)

passengers %>%
  filter(airline %in% top_5) %>%
  group_by(date, airline) %>%
  summarise(total_passengers = sum(passengers, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = date, y = total_passengers, linetype = airline)) +
  geom_line(linewidth = 1) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Monthly Passenger Trends for Top 5 Airlines",
    x = "Date",
    y = "Passengers"
  ) +
  theme_minimal()
```

## Use airport metric table

### Example: Total Passengers metric from metrics file
```{r}
total_metric <- metrics %>%
  filter(metric == "Total Passengers") %>%
  arrange(date)

print(total_metric)

ggplot(total_metric, aes(x = date, y = value)) +
  geom_line(linewidth = 1) +
  geom_point() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Official Total Passenger Metric Over Time",
    x = "Date",
    y = "Passengers"
  ) +
  theme_minimal()
```

## Create a baseline model

### Simple time trend model using monthly totals

```{r}
baseline_data <- monthly_total %>%
  mutate(
    time_index = row_number(),
    month_factor = factor(month_num)
  )

baseline_model <- lm(total_passengers ~ time_index + month_factor, data = baseline_data)
summary(baseline_model)

baseline_data$predicted_baseline <- predict(baseline_model, newdata = baseline_data)

ggplot(baseline_data, aes(x = date)) +
  geom_line(aes(y = total_passengers, linetype = "Actual"), linewidth = 1) +
  geom_line(aes(y = predicted_baseline, linetype = "Baseline Prediction"), linewidth = 1) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Actual vs Baseline Predicted Passengers",
    x = "Date",
    y = "Passengers",
    linetype = ""
  ) +
  theme_minimal()

```

## AO4 Simulation Framework

Since your current data does not yet include:
- sentiment
- engagement
- behavior
we create proxy scenario variables manually.

```{r}
simulation_data <- baseline_data %>%
  mutate(
    # Placeholder proxy inputs (edit these later)
    engagement_score = 50,   # 0 to 100
    sentiment_score = 0.10,  # -1 to +1
    behavior_score = 0.20    # 0 to 1
  )
```

## Example simulation formula:

estimated uplift = engagement effect + sentiment effect + behavior effect

```{r}
simulation_data <- simulation_data %>%
  mutate(
    estimated_uplift_pct =
      (engagement_score * 0.0015) +
      (sentiment_score * 0.08) +
      (behavior_score * 0.12),

    simulated_passengers = predicted_baseline * (1 + estimated_uplift_pct)
  )

print(simulation_data %>%
        select(date, total_passengers, predicted_baseline, estimated_uplift_pct, simulated_passengers))

ggplot(simulation_data, aes(x = date)) +
  geom_line(aes(y = predicted_baseline, linetype = "Baseline"), linewidth = 1) +
  geom_line(aes(y = simulated_passengers, linetype = "Simulated Strategy Impact"), linewidth = 1) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Simulation of Future Community Engagement Impact",
    x = "Date",
    y = "Passengers",
    linetype = ""
  ) +
  theme_minimal()
```

## Scenario testing

Best case, moderate case, conservative case

```{r}
scenarios <- tibble(
  scenario = c("Conservative", "Moderate", "Best Case"),
  engagement_score = c(40, 60, 80),
  sentiment_score = c(0.05, 0.12, 0.20),
  behavior_score = c(0.10, 0.20, 0.35)
)

scenario_results <- scenarios %>%
  mutate(
    estimated_uplift_pct =
      (engagement_score * 0.0015) +
      (sentiment_score * 0.08) +
      (behavior_score * 0.12)
  ) %>%
  crossing(baseline_data %>% select(date, predicted_baseline)) %>%
  mutate(
    simulated_passengers = predicted_baseline * (1 + estimated_uplift_pct)
  )

print(scenario_results)

ggplot(scenario_results, aes(x = date, y = simulated_passengers, linetype = scenario)) +
  geom_line(linewidth = 1) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Scenario Simulation for Community Engagement Strategies",
    x = "Date",
    y = "Simulated Passengers",
    linetype = "Scenario"
  ) +
  theme_minimal()
```

## Simple ROI proxy framework

Since real revenue is not in the dataset, create a proxy value per passenger
Adjust this assumption based on your project logic

```{r}
value_per_passenger <- 18

scenario_roi <- scenario_results %>%
  mutate(
    baseline_value = predicted_baseline * value_per_passenger,
    simulated_value = simulated_passengers * value_per_passenger,
    incremental_value = simulated_value - baseline_value
  ) %>%
  group_by(scenario) %>%
  summarise(
    total_baseline_value = sum(baseline_value, na.rm = TRUE),
    total_simulated_value = sum(simulated_value, na.rm = TRUE),
    total_incremental_value = sum(incremental_value, na.rm = TRUE),
    .groups = "drop"
  )

print(scenario_roi)
```




```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(scales)

data <- read_excel("ao4_model_ready_dataset.xlsx")

data <- data %>%
  arrange(year, month) %>%
  mutate(
    date = as.Date(paste(year, month, "01", sep = "-"))
  )
```


## Passenger trend over time

```{r}
ggplot(data, aes(x = date, y = total_passengers)) +
  geom_line(linewidth = 1) +
  geom_point() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Monthly Passenger Trend",
    x = "Date",
    y = "Passengers"
  ) +
  theme_minimal()
```

## Engagement (events) over time

```{r}
ggplot(data, aes(x = date, y = total_events)) +
  geom_line(linewidth = 1) +
  geom_point() +
  labs(
    title = "Community Engagement Activity Over Time",
    x = "Date",
    y = "Number of Events"
  ) +
  theme_minimal()
```

## MOST IMPORTANT GRAPH (AO4)


```{r}
ggplot(data, aes(x = total_events, y = total_passengers)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Impact of Community Engagement on Passenger Volume",
    x = "Number of Events",
    y = "Passengers"
  ) +
  theme_minimal()
```

## Combined dual-axis style (visual storytelling)

```{r}
ggplot(data, aes(x = date)) +
  geom_line(aes(y = total_passengers, linetype = "Passengers"), linewidth = 1) +
  geom_line(aes(y = total_events * 10000, linetype = "Events (scaled)"), linewidth = 1) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Passengers vs Engagement Activity",
    x = "Date",
    y = "Passengers",
    linetype = ""
  ) +
  theme_minimal()
```

## Model prediction visualization

```{r}
data <- data %>%
  mutate(
    time_index = row_number(),
    month_factor = factor(month)
  )

model <- lm(total_passengers ~ total_events + time_index + month_factor, data = data)

data$predicted <- predict(model, newdata = data)

ggplot(data, aes(x = date)) +
  geom_line(aes(y = total_passengers, linetype = "Actual"), linewidth = 1) +
  geom_line(aes(y = predicted, linetype = "Predicted"), linewidth = 1) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Actual vs Predicted Passengers",
    x = "Date",
    y = "Passengers",
    linetype = ""
  ) +
  theme_minimal()
```

## Simulation graph (AO4 highlight)


```{r}
library(dplyr)
library(ggplot2)
library(scales)

data <- data %>%
  mutate(
    simulated_high = predicted * 1.10,
    simulated_low = predicted * 0.95
  )

ggplot(data, aes(x = date)) +
  geom_line(aes(y = predicted, color = "Baseline", linetype = "Baseline"), linewidth = 1) +
  geom_line(aes(y = simulated_high, color = "High Engagement", linetype = "High Engagement"), linewidth = 1) +
  geom_line(aes(y = simulated_low, color = "Low Engagement", linetype = "Low Engagement"), linewidth = 1) +
  scale_color_manual(values = c(
    "Baseline" = "black",
    "High Engagement" = "blue",
    "Low Engagement" = "red"
  )) +
  scale_linetype_manual(values = c(
    "Baseline" = "solid",
    "High Engagement" = "dashed",
    "Low Engagement" = "dotted"
  )) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Simulation of Engagement Strategies",
    x = "Date",
    y = "Passengers",
    color = "Scenario",
    linetype = "Scenario"
  ) +
  theme_minimal()
```

# AO4 - CE Cities Engagement Analysis

```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(scales)
library(stringr)
library(tidyr)
```

```{r}
ce <- read_excel("FY 24-25 CE Cities-Engagements.xlsx", sheet = "Sheet1")
```


```{r}
names(ce) <- c(
  "city_name",
  "county",
  "total_visits",
  "total_engagements",
  "total_presentations",
  "total_community_booths",
  "leads"
)

```

```{r}
ce <- ce %>%
  mutate(
    city_name = str_trim(city_name),
    county = str_trim(county),
    leads = str_trim(leads),
    across(c(total_visits, total_engagements, total_presentations, total_community_booths), as.numeric)
  )
```

## Inspect structure

```{r}
glimpse(ce)
summary(ce)
```

## Overall summary

```{r}
overall_summary <- ce %>%
  summarise(
    total_cities = n(),
    total_visits = sum(total_visits, na.rm = TRUE),
    total_engagements = sum(total_engagements, na.rm = TRUE),
    total_presentations = sum(total_presentations, na.rm = TRUE),
    total_booths = sum(total_community_booths, na.rm = TRUE),
    avg_engagements_per_city = mean(total_engagements, na.rm = TRUE)
  )

print(overall_summary)
```

## Top cities by total engagements

```{r}
top_cities <- ce %>%
  arrange(desc(total_engagements)) %>%
  select(city_name, county, total_engagements, total_visits, total_presentations, total_community_booths) %>%
  slice_head(n = 10)

print(top_cities)
```

## County-level summary

```{r}
county_summary <- ce %>%
  group_by(county) %>%
  summarise(
    cities = n(),
    total_visits = sum(total_visits, na.rm = TRUE),
    total_engagements = sum(total_engagements, na.rm = TRUE),
    total_presentations = sum(total_presentations, na.rm = TRUE),
    total_booths = sum(total_community_booths, na.rm = TRUE),
    avg_engagements_per_city = mean(total_engagements, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(total_engagements))

print(county_summary)
```

## Lead-level summary

```{r}
lead_summary <- ce %>%
  mutate(leads = ifelse(is.na(leads) | leads == "", "Unknown", leads)) %>%
  group_by(leads) %>%
  summarise(
    cities = n(),
    total_visits = sum(total_visits, na.rm = TRUE),
    total_engagements = sum(total_engagements, na.rm = TRUE),
    total_presentations = sum(total_presentations, na.rm = TRUE),
    total_booths = sum(total_community_booths, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(total_engagements))

print(lead_summary)
```

## Create efficiency metrics

```{r}
ce <- ce %>%
  mutate(
    engagements_per_visit = ifelse(total_visits > 0, total_engagements / total_visits, NA),
    presentations_per_visit = ifelse(total_visits > 0, total_presentations / total_visits, NA),
    booths_per_visit = ifelse(total_visits > 0, total_community_booths / total_visits, NA)
  )

efficiency_summary <- ce %>%
  arrange(desc(engagements_per_visit)) %>%
  select(city_name, county, total_visits, total_engagements, engagements_per_visit) %>%
  slice_head(n = 10)

print(efficiency_summary)
```

## Create a simple weighted engagement score

```{r}
ce <- ce %>%
  mutate(
    engagement_score =
      (total_engagements * 1.0) +
      (total_presentations * 50) +
      (total_community_booths * 40) +
      (total_visits * 10)
  )

score_summary <- ce %>%
  arrange(desc(engagement_score)) %>%
  select(city_name, county, engagement_score, total_engagements, total_visits) %>%
  slice_head(n = 10)

print(score_summary)
```

## GRAPHS 
## Top 10 cities by total engagements

```{r}
ggplot(top_cities, aes(x = reorder(city_name, total_engagements), y = total_engagements)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Top 10 Cities by Total Engagements",
    x = "City",
    y = "Total Engagements"
  ) +
  theme_minimal()
```

## County comparison

```{r}
ggplot(county_summary, aes(x = reorder(county, total_engagements), y = total_engagements)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Total Engagements by County",
    x = "County",
    y = "Total Engagements"
  ) +
  theme_minimal()
```

## Visits vs engagements relationship

```{r}
ggplot(ce, aes(x = total_visits, y = total_engagements)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Relationship Between Visits and Engagements",
    x = "Total Visits",
    y = "Total Engagements"
  ) +
  theme_minimal()
```

## Top cities by engagement score

```{r}
top_score <- ce %>%
  arrange(desc(engagement_score)) %>%
  slice_head(n = 10)

ggplot(top_score, aes(x = reorder(city_name, engagement_score), y = engagement_score)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Top 10 Cities by Weighted Engagement Score",
    x = "City",
    y = "Engagement Score"
  ) +
  theme_minimal()
```

## Lead comparison

```{r}
ggplot(lead_summary, aes(x = reorder(leads, total_engagements), y = total_engagements)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Total Engagements by Lead",
    x = "Lead",
    y = "Total Engagements"
  ) +
  theme_minimal()
```

# Monthly Engagement 

## Total engagement across all cities

```{r}
total_engagement <- ce %>%
  summarise(
    total_engagements = sum(total_engagements, na.rm = TRUE),
    total_visits = sum(total_visits, na.rm = TRUE),
    total_presentations = sum(total_presentations, na.rm = TRUE),
    total_booths = sum(total_community_booths, na.rm = TRUE)
  )

print(total_engagement)
```

## Convert to monthly averages (12 months assumption)

```{r}
monthly_engagement <- total_engagement %>%
  mutate(
    monthly_engagements = total_engagements / 12,
    monthly_visits = total_visits / 12,
    monthly_presentations = total_presentations / 12,
    monthly_booths = total_booths / 12
  )

print(monthly_engagement)
```

```{r}
months <- data.frame(
  month = 1:12,
  year = 2025
)

engagement_time <- months %>%
  mutate(
    total_events = monthly_engagement$monthly_engagements,
    visits = monthly_engagement$monthly_visits,
    presentations = monthly_engagement$monthly_presentations,
    booths = monthly_engagement$monthly_booths
  )

print(engagement_time)
```

```{r}
library(readxl)

passengers <- read_excel("ao4_model_ready_dataset.xlsx")

data <- passengers %>%
  left_join(engagement_time, by = c("year", "month"))

head(data)
```




```{r}
library(readr)
library(dplyr)
library(lubridate)

profile <- read_csv("Profile Performance.csv", show_col_types = FALSE)

colnames(profile)
head(profile)
```


```{r}
library(dplyr)
library(lubridate)

profile <- profile %>%
  mutate(
    date = as.Date(Date, format = "%m-%d-%Y"),  # <-- FIXED FORMAT
    year = year(date),
    month = month(date),

    impressions = as.numeric(Impressions),
    engagement = as.numeric(Engagements),
    comments = as.numeric(Comments),
    shares = as.numeric(Shares)
  )
```



