---
title: "Social Media Profile VS Passengers"
author: "Saurabh Parate"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
    code-fold: show
    code-summary: "Show code"
    df-print: paged
execute:
  echo: true
  warning: false
  message: false
  error: false
---

# Data

```{r}
packages <- c("readr", "dplyr", "lubridate", "ggplot2", "scales", "tidyr")
installed <- rownames(installed.packages())

for (p in packages) {
  if (!(p %in% installed)) install.packages(p)
}

library(readr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(scales)
library(tidyr)
```

```{r}
profile <- read_csv("Profile Performance.csv", show_col_types = FALSE)
passengers <- read_csv("pnfs_passengers_long.csv", show_col_types = FALSE)

# 3) Clean passenger data
passengers_monthly <- passengers %>%
  mutate(
    date = as.Date(date),
    year = year(date),
    month = month(date)
  ) %>%
  filter(year == 2025) %>%
  group_by(year, month) %>%
  summarise(
    total_passengers = sum(passengers, na.rm = TRUE),
    .groups = "drop"
  )

print(passengers_monthly)
```


```{r}
profile_clean <- profile %>%
  mutate(
    date = mdy(Date),

    Audience = as.numeric(gsub(",", "", Audience)),
    `Net Audience Growth` = as.numeric(gsub(",", "", `Net Audience Growth`)),
    `Audience Gained` = as.numeric(gsub(",", "", `Audience Gained`)),
    `Published Posts (Total)` = as.numeric(gsub(",", "", `Published Posts (Total)`)),
    Impressions = as.numeric(gsub(",", "", Impressions)),
    `Video Views` = as.numeric(gsub(",", "", `Video Views`)),
    Engagements = as.numeric(gsub(",", "", Engagements)),
    Reactions = as.numeric(gsub(",", "", Reactions)),
    Comments = as.numeric(gsub(",", "", Comments)),
    Shares = as.numeric(gsub(",", "", Shares))
  ) %>%
  mutate(
    year = year(date),
    month = month(date)
  ) %>%
  filter(year == 2025)
```

# Aggregate profile data to monthly level

```{r}
profile_monthly <- profile_clean %>%
  group_by(year, month) %>%
  summarise(
    total_impressions = sum(Impressions, na.rm = TRUE),
    total_engagements = sum(Engagements, na.rm = TRUE),
    total_posts = sum(`Published Posts (Total)`, na.rm = TRUE),
    avg_audience = mean(Audience, na.rm = TRUE),
    net_audience_growth = sum(`Net Audience Growth`, na.rm = TRUE),
    total_video_views = sum(`Video Views`, na.rm = TRUE),
    total_reactions = sum(Reactions, na.rm = TRUE),
    total_comments = sum(Comments, na.rm = TRUE),
    total_shares = sum(Shares, na.rm = TRUE),
    .groups = "drop"
  )

print(profile_monthly)
```

# Merge with passenger data

```{r}
data_2025 <- passengers_monthly %>%
  left_join(profile_monthly, by = c("year", "month")) %>%
  arrange(year, month) %>%
  mutate(
    date = as.Date(sprintf("%d-%02d-01", year, month))
  )

print(data_2025)
```


#Check missing values

```{r}
colSums(is.na(data_2025))
```

```{r}
data_2025 <- data_2025 %>%
  mutate(
    across(
      c(total_impressions, total_engagements, total_posts, net_audience_growth,
        total_video_views, total_reactions, total_comments, total_shares),
      ~replace_na(., 0)
    )
  )
```

# Correlation check


```{r}
cor_vars <- data_2025 %>%
  select(
    total_passengers,
    total_impressions,
    total_engagements,
    total_posts,
    avg_audience,
    net_audience_growth,
    total_video_views
  )

print(cor(cor_vars, use = "complete.obs"))
```

# Simple comparison models


```{r}
model1 <- lm(total_passengers ~ total_engagements, data = data_2025)
summary(model1)

model2 <- lm(total_passengers ~ total_impressions, data = data_2025)
summary(model2)

model3 <- lm(total_passengers ~ total_posts, data = data_2025)
summary(model3)
```

# Stronger model


```{r}
data_2025 <- data_2025 %>%
  mutate(
    time_index = row_number(),
    month_factor = factor(month)
  )

model_final <- lm(
  total_passengers ~ total_engagements + total_impressions + total_posts + time_index,
  data = data_2025
)

summary(model_final)
```


# Predicted values

```{r}
data_2025 <- data_2025 %>%
  mutate(
    predicted_passengers = predict(model_final, newdata = data_2025)
  )

```

# Graphs
## Passenger trend

```{r}
ggplot(data_2025, aes(x = date, y = total_passengers)) +
  geom_line(linewidth = 1) +
  geom_point() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Monthly Passenger Volume (2025)",
    x = "Month",
    y = "Passengers"
  ) +
  theme_minimal()
```

## Profile engagements trend

```{r}
ggplot(data_2025, aes(x = date, y = total_engagements)) +
  geom_line(linewidth = 1) +
  geom_point() +
  labs(
    title = "Monthly Profile Engagements (2025)",
    x = "Month",
    y = "Engagements"
  ) +
  theme_minimal()
```

## Engagements vs Passengers

```{r}
ggplot(data_2025, aes(x = total_engagements, y = total_passengers)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Profile Engagements vs Passenger Volume",
    x = "Total Engagements",
    y = "Passengers"
  ) +
  theme_minimal()
```

## Impressions vs Passengers

```{r}
ggplot(data_2025, aes(x = total_impressions, y = total_passengers)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Profile Impressions vs Passenger Volume",
    x = "Total Impressions",
    y = "Passengers"
  ) +
  theme_minimal()
```

## Actual vs Predicted


```{r}
ggplot(data_2025, aes(x = date)) +
  geom_line(aes(y = total_passengers, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = predicted_passengers, color = "Predicted"), linewidth = 1) +
  scale_y_continuous(labels = comma) +
  scale_color_manual(values = c("Actual" = "black", "Predicted" = "blue")) +
  labs(
    title = "Actual vs Predicted Passengers",
    x = "Month",
    y = "Passengers",
    color = ""
  ) +
  theme_minimal()
```

# Simulation scenarios


```{r}
# Create scenarios based on engagement changes
simulation <- data_2025 %>%
  mutate(
    low_engagement = total_engagements * 0.8,
    high_engagement = total_engagements * 1.2
  )

# Predict passengers under each scenario
simulation <- simulation %>%
  mutate(
    baseline_pred = predict(model_final, newdata = data_2025),

    low_pred = predict(model_final,
      newdata = data_2025 %>%
        mutate(total_engagements = low_engagement)
    ),

    high_pred = predict(model_final,
      newdata = data_2025 %>%
        mutate(total_engagements = high_engagement)
    )
  )
```

# Visualize simulation (KEY GRAPH)

```{r}
ggplot(simulation, aes(x = date)) +
  geom_line(aes(y = baseline_pred, color = "Baseline"), linewidth = 1) +
  geom_line(aes(y = low_pred, color = "Low Engagement"), linewidth = 1) +
  geom_line(aes(y = high_pred, color = "High Engagement"), linewidth = 1) +
  scale_y_continuous(labels = comma) +
  scale_color_manual(values = c(
    "Baseline" = "black",
    "Low Engagement" = "red",
    "High Engagement" = "blue"
  )) +
  labs(
    title = "Simulation of Engagement Impact on Passenger Volume",
    x = "Month",
    y = "Passengers",
    color = "Scenario"
  ) +
  theme_minimal()
```

# Convert to ROI

proxy value per passenger

```{r}
 value_per_passenger <- 20
```

A simulation framework was developed to estimate the impact of changes in engagement levels on passenger volume. By adjusting engagement inputs and predicting passenger outcomes, the model provides a forward-looking estimate of potential ROI from community engagement strategies.

```{r}
roi_results <- simulation %>%
  mutate(
    baseline_value = baseline_pred * value_per_passenger,
    high_value = high_pred * value_per_passenger,
    incremental_value = high_value - baseline_value
  ) %>%
  summarise(
    total_incremental_value = sum(incremental_value, na.rm = TRUE)
  )

print(roi_results)
```

A 20% increase in engagement is associated with approximately $5.7 million in incremental passenger value over the year

Under a high-engagement scenario (+20%), the model estimates an incremental value of approximately $5.7 million, based on a conservative proxy of $20 per passenger. This reflects the potential economic impact of increased community engagement efforts.

# Where engagement matters most (seasonally)

```{r}
ggplot(simulation, aes(x = date)) +
  geom_line(aes(y = high_pred - baseline_pred), linewidth = 1) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Incremental Passengers from Increased Engagement",
    x = "Month",
    y = "Additional Passengers"
  ) +
  theme_minimal()
```

